Enhance Road Detection Data Processing of LiDAR Point Clouds to Specifically Identify Unmarked Gravel Rural Roads

### 2023-05-25 Huston_Paper.tex
More Updates

			{Scanning LiDAR training data requires extraction from manually defined gravel, asphalt, and "unknown" terrain areas on aggregated point clouds. Aggregating LiDAR data into a single point cloud was completed by post processing rosbags that contained scanning LiDAR, GPS, and IMU data of an unmarked road. Scanning LiDAR timestamps were matched to the closest GNSS and IMU timestamp data. Closest matching GNSS and IMU data informed the location and orientation of the LiDAR scan. Point cloud translation and rotation was accomplished using transformation matrices derived from GPS and IMU data. Rotation matrices were created using extracted roll, pitch, and yaw data from the IMU. LiDAR origin was found using GPS longitude, latitude, and altitude data. Physical distances between the GPS, IMU, and LiDAR sensors was rectified by obtaining reference frames provided by AutonomouStuff. GPS coordinates were offset by the current orientation and converted the ground truth to the LiDAR frame. GPS, IMU, and LiDAR reference frames and rotational updates were combined into trajectory vectors. Consecutive point cloud scans may then translated and rotated with derived transformation matrices (Fig. \ref{fig:road_areas_annotated_22}). While not as robust as more sophisticated methods of point cloud aggregation such as NDT or ICP scan matching, over shorter distances using this method proved to be adequate for this work.}

			\begin{figure}[H]
				\centering
				\includegraphics[width=0.75\linewidth]{figures/classified_point_cloud_example_redux}
				\caption[Aggregated Point Cloud Data]{Point cloud of scanning LiDAR data using GPS and IMU data to determine point of origin and orientation for the LiDAR was used to create a combined point cloud map.}
			\end{figure}

@@

			{MATLAB's Classification Learner Application was used to generate a Random Decision Forests for each considered scanning LiDAR channel. Decision tree depth, number of learners, and number of features to consider for each binary split are hyper-parameters that were optimized using MATLAB's Bayesian Optimization. Bayesian Optimization automates the manual hyper-parameter tuning process by creating a surrogate model representing the objective function, in this the upper confidence interval of the classification error objective model (Fig. \ref{fig:c2_min_class_error}). Initialization of the objective function is completed by training a number of RDFs using randomly selected hyper-parameters. Out-of-bag error is then found and used to create a series of Gaussian curves. Additional hyper-parameter values are chosen by an acquisition function to determine their utility in minimizing out-of-bag error. After a set number of iterations, best hyper-parameters are chosen based minimum upper confidence interval of the classification error objective model. Hyperparameter tuning may lead to over-fitting to the training data base. Testing for over-fitting was accomplished by comparing the out-of-bag training error to the validation error (Fig. \ref{fig:train_vs_valid_overfit_test2}). Confusion matrices provide classification accuracy information on validation database by classifying each data set in the validation database and comparing actual and predicted classes, resulting in a matrix that visualizes the accuracy for each terrain type as well as which terrain types are more likely to be confused for a different class (Fig. \ref{fig:vali_err_conf_mat}). This process was completed for each of the three considered scanning LiDAR channels used for classification.}

@@

			{Classifying consecutive LiDAR scans was accomplished by examination of the nine arcs of interest in front and forty-five degree angles left and right from the front of the experimental apparatus (Fig. \ref{fig:area_example}). Features from each arc were extracted and the arc was classified (Fig. \ref{fig:raw_classification_results} shows the average spatial coordinates of each classified arc). Previously-created manually defined gravel, asphalt, and side-of-road truth areas (denoted as Testing areas in Fig. \ref{fig:test_vs_train_areas})were projected onto the classified point cloud for the determination of classification accuracy . Averaged x, y, and z coordinates were taken of each arc and projected unto a manually defined truth area map (Fig. \ref{fig:rm_db_4_toc}). Accuracy scores was calculated based on the distribution of classified points with each area, allowing evaluation of unmarked road detection performance. Classified point clouds were visually inspected to determine if the manual detection of a road surface may be accomplished.  If consecutive points from each channel in any direction had two or more matching gravel or asphalt classification and was evenly spaced, a gravel or road surface area was projected. Guessed road surface areas were then compared to actual areas to determine the method's ability at unmarked gravel road discovery, indicating the method's efficiency at detecting intercepting gravel roads. }

@@

		{Classification of real-world data outside of the training data set presents difficulties to the generated Random Decision Forest Classifiers generated in this work. Due to the side of the road was being re-grassed near a portion of the asphalt road, the terrain classification algorithm struggled to differentiate the dirt from gravel, leading to some miss-classification of the side-of-road areas. Examination of the feature distribution shows a high amount of overlap (Fig. \ref{fig:dirt_v_gravel}). Exploitation of confidence scores to adjust results by requiring a very high confidence for positive gravel surface classification was explored. All gravel surfaces that did not meet a 80\% confidence score was re-labeled as "unknown", however these adjustment had either a negligible or detrimental impact on the final results. Higher confidence score requirements resulted in "unknown" over-classification, while too low resulted in "gravel" over-classification. Overall it was found that an average true-positive accuracy of all intercepting gravel surfaces was 67.54\%. If any asphalt miss-classification on the gravel areas is included, the average true-positive road detection accuracy increases to 79.18\%. It was found that an average true-positive accuracy of all asphalt surfaces was 78.06\%. Differentiation between gravel road surfaces and surrounding terrain is necessary to determine boundaries for trajectory tracking and navigation. Side-of-road areas were established (Fig. \ref{fig:prepostadjust}), and the overall average for "unknown" terrain type was found to be 77.94\%.}

@@

		% Summary of Work & Closing Statements
		{High definition scanning LiDAR and GPS data was gathered on Blackburn Road and an intercepting gravel lot in Athens Ohio. Properties describing surface roughness and remission data were extracted and used to train Random Decision Forests. Terrain classification of consecutive scans compared to manually detected road surface areas yielded accuracy scores for gauging algorithm capability for gravel surface detection. It was found that a true-positive accuracy for gravel and asphalt surfaces was 67.54\% and 87.05\% respectively at an estimated rate of 13.19 milliseconds per 360 degree scan. Classified point clouds were examined and gravel areas were projected and compared with actual gravel surface areas. Overlapping results between detected and actual road surface areas resulted in 91.67\% intercepting gravel road detection accuracy. In conjunction with the overlapping areas and the high true-positive detection accuracy of gravel surfaces, this method of gravel road detection is promising for future exploration of additional roads and road surfaces. Work as described addressed the problem of road surface detection on unmarked gravel and chipseal roads. The described work was a method of detecting physical unmarked gravel and chipseal roads by using a terrain classification approach to predicting road surface area. The impact of this work is that autonomous vehicles using the developed method may be able to detect unmarked gravel and chipseal road surfaces, allowing autonomous operations on 1.5 million miles of previously undetected rural roads.}

>>>>

			{Scanning LiDAR training data requires extraction from manually defined gravel, asphalt, and "unknown" terrain areas on aggregated point clouds. Aggregating LiDAR data into a single point cloud was completed by post processing rosbags that contained scanning LiDAR, GPS, and IMU data of an unmarked road. Scanning LiDAR timestamps were matched to the closest GNSS and IMU timestamp data. Closest matching GNSS and IMU data informed the location and orientation of the LiDAR scan. Point cloud translation and rotation was accomplished using transformation matrices derived from GPS and IMU data. Rotation matrices were created using extracted roll, pitch, and yaw data from the IMU. LiDAR origin was found using GPS longitude, latitude, and altitude data. Physical distances between the GPS, IMU, and LiDAR sensors was rectified by obtaining reference frames provided by AutonomouStuff. GPS coordinates were offset by the current orientation and converted the ground truth to the LiDAR frame. GPS, IMU, and LiDAR reference frames and rotational updates were combined into trajectory vectors. Consecutive point cloud scans may then translated and rotated with derived transformation matrices (Fig. \ref{fig:pc_example}). While not as robust as more sophisticated methods of point cloud aggregation such as NDT or ICP scan matching, over shorter distances using this method proved to be adequate for this work.}

			\begin{figure}[H]
				\centering
				\includegraphics[width=0.75\linewidth]{figures/pc_example}
				\caption[Aggregated Point Cloud Data]{Point cloud of scanning LiDAR data using GPS and IMU data to determine point of origin and orientation for the LiDAR was used to create a combined point cloud map.}
				\label{fig:pc_example}
			\end{figure}

@@

			{MATLAB's Classification Learner Application was used to generate a Random Decision Forests for each considered scanning LiDAR channel. Decision tree depth, number of learners, and number of features to consider for each binary split are hyper-parameters that were optimized using MATLAB's Bayesian Optimization. Bayesian Optimization automates the manual hyper-parameter tuning process by creating a surrogate model representing the objective function, in this the upper confidence interval of the classification error objective model (Fig. \ref{fig:c2_min_class_error}). Initialization of the objective function is completed by training a number of RDFs using randomly selected hyper-parameters. Out-of-bag error is then found and used to create a series of Gaussian curves. Additional hyper-parameter values are chosen by an acquisition function to determine their utility in minimizing out-of-bag error. After a set number of iterations, best hyper-parameters are chosen based minimum upper confidence interval of the classification error objective model. Hyperparameter tuning may lead to over-fitting to the training data base. Testing for over-fitting was accomplished by comparing the out-of-bag training error to the validation error (Fig. \ref{fig:train_vs_valid_overfit_test2}). Confusion matrices provide validation database classification accuracy information by classifying each data set in the validation database and comparing actual and predicted classes, resulting in a matrix that visualizes the accuracy for each terrain type as well as which terrain types are more likely to be confused for a different class (Fig. \ref{fig:vali_err_conf_mat}). This process was completed for each of the three considered scanning LiDAR channels used for classification.}

@@

			{Classifying consecutive LiDAR scans was accomplished by examination of the nine arcs of interest in front and forty-five degree angles left and right from the front of the experimental apparatus (Fig. \ref{fig:area_example}). Features from each arc were extracted and the arc was classified (Fig. \ref{fig:raw_classification_results} shows the average spatial coordinates of each classified arc). Previously-created manually defined gravel, asphalt, and side-of-road truth areas (denoted as Testing areas in Fig. \ref{fig:test_vs_train_areas}) were projected onto the classified point cloud for the determination of classification accuracy. Averaged x, y, and z coordinates were taken of each arc and projected unto a manually defined truth area map (Fig. \ref{fig:rm_db_4_toc}). Accuracy scores was calculated based on the distribution of classified points with each area, allowing evaluation of unmarked road detection performance. Classified point clouds were visually inspected to determine if the manual detection of a road surface may be accomplished. If consecutive points from each channel in any direction had two or more matching gravel or asphalt classification and was evenly spaced, a gravel or road surface area was projected. Guessed road surface areas were then compared to actual areas to determine the method's ability at unmarked gravel road discovery, indicating the method's efficiency at detecting intercepting gravel roads. }

@@

		{Classification of real-world data outside of the training data set presents difficulties to the generated Random Decision Forest Classifiers generated in this work. Due to the side of the road was being re-grassed near a portion of the asphalt road, the terrain classification algorithm struggled to differentiate the dirt from gravel, leading to some miss-classification of the side-of-road areas. Examination of the feature distribution shows a high amount of overlap (Fig. \ref{fig:dirt_v_gravel}). Exploitation of confidence scores to adjust results by requiring a very high confidence for positive gravel surface classification was explored. All gravel surfaces that did not meet a 80\% confidence score was re-labeled as "unknown", however these adjustment had either a negligible or detrimental impact on the final results. Higher confidence score requirements resulted in "unknown" over-classification, while too low resulted in "gravel" over-classification. Overall it was found that an average true-positive accuracy of all intercepting gravel surfaces was 67.54\% (Fig. \ref{fig:prepostadjust}). If any asphalt miss-classification on the gravel areas is included, the average true-positive road detection accuracy increases to 79.18\%. It was found that an average true-positive accuracy of all asphalt surfaces was 78.06\%. Differentiation between gravel road surfaces and surrounding terrain is necessary to determine boundaries for trajectory tracking and navigation. Overall average distribution for "unknown" terrain class within side-of-road areas was found to be 77.94\%.}

@@

		% Summary of Work & Closing Statements
		{High definition scanning LiDAR and GPS data was gathered on Blackburn Road and intercepting gravel lots in Athens Ohio. Properties describing surface roughness and remission data were extracted and used to train Random Decision Forests. Terrain classification of consecutive scans compared to manually detected road surface areas yielded accuracy scores for gauging algorithm capability for gravel surface detection. It was found that a true-positive accuracy for gravel and asphalt surfaces was 67.54\% and 87.05\% respectively at an estimated rate of 13.19 milliseconds per 360 degree scan. Classified point clouds were examined and gravel areas were projected and compared with actual gravel surface areas. Overlapping results between detected and actual road surface areas resulted in 91.67\% intercepting gravel road detection accuracy. In conjunction with the overlapping areas and the high true-positive detection accuracy of gravel surfaces, this method of gravel road detection is promising for future exploration of additional roads and road surfaces. Work as described addressed the problem of road surface detection on unmarked gravel and chipseal roads. The described work was a method of detecting physical unmarked gravel and chipseal roads by using a terrain classification approach to predicting road surface area. The impact of this work is that autonomous vehicles using the developed method may be able to detect unmarked gravel and chipseal road surfaces, allowing autonomous operations on 1.5 million miles of previously undetected rural roads.}
